Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-19 forecast/paper/paper.tex
Copyedit abstract

% At least three keywords are required at submission. Please provide three to
% five keywords, separated by the pipe symbol. 
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Reliable short-term, location-specific forecasts of a pandemic are a
key ingredient in effective public health decision making. Throughout the
COVID-19 pandemic, a plethora of real-time data signals were
collected, curated, and made publicly available by our team and data
partners, providing multiple views of the pandemic's progress. The
goal of this work is to measure the practical utility of these
auxiliary indicators from a forecasting perspective. We focus on five
indicators, derived from medical insurance claims data, web search
queries, and survey responses. For each indicator, we ask whether its
inclusion in a simple model leads to improved forecasting
performance relative to a similar model excluding it. Performance is considered across two tasks
at a forecasting horizon of up to about one month:
probabilistic forecasting of seven-day average case rates and binary
prediction of ``hotspots.'' Since the values of indicators are
commonly revised over time, we take special care to ensure that the data provided
to a forecaster is the version that would have been available at the
time that the forecast was made. We demonstrate how less careful
backtesting can lead to misleading retrospective evaluations. Our analysis
suggests consistent, although 
modest, gains in forecasting performance by using these indicators,
and further investigation suggests that these gains are related to periods when the
auxiliary signals are acting as ``leading 
indicators'' of case rates. 
\attn{Evaluating forecasters thoroughly and fairly is an important way that
we can learn from this pandemic to do better in the next. 251 :(}

% Please provide an abstract of no more than 250 words in a single paragraph.
% Abstracts should explain to the general reader the major contributions of the
% article. References in the abstract must be cited in full within the abstract
% itself and cited in the text.
\end{abstract}

>>>>

% At least three keywords are required at submission. Please provide three to
% five keywords, separated by the pipe symbol. 
\keywords{COVID-19 $|$ forecasting $|$ hotspot prediction $|$ time series | 
  digital surveillance}

\begin{abstract}
  Reliable, short-term forecasts of traditional public health reporting streams
  (such as cases, hospitalizations, and deaths) are a key ingredient in
  effective public health decision-making during a pandemic. Since April 2020,
  our research group has worked with data partners to collect, curate, and make
  publicly available numerous real-time COVID-19 indicators, providing multiple
  views of pandemic activity.  This paper studies the utility of these
  indicators from a forecasting perspective. We focus on five indicators,
  derived from medical insurance claims data, web search queries, and online
  survey responses. For each indicator, we ask whether its inclusion in a simple
  model leads to improved predictive accuracy relative to a similar model
  excluding it. We consider both probabilistic forecasting of confirmed COVID-19 
  case rates and binary prediction of case ``hotspots''.  Since the values of 
  indicators (and case rates) are commonly revised over time, we take special
  care to ensure that the data provided to a forecaster is the version that
  would have been available at the time the forecast was made.   
  % We demonstrate how less careful backtesting can lead to misleading
  % retrospective evaluations.   
  Our  analysis shows that consistent but modest gains in predictive accuracy 
  are obtained by using these indicators, and furthermore, these gains are
  related to periods in which the auxiliary indicators behave as ``leading
  indicators'' of case rates.     
\end{abstract}
